Self-interacting dark matter model without dark energy in cosmology

### 2024-12-22 article/article.tex
use getdist to draw the mcmc result

\subsection{Model 1: relation to the LambdaCDM model}

@@

   To constrain the relevant IDM models (\cite{refId0}), we use the newly revised
   observational $H(z)$ data (OHD)(\cite{Zhang_2014};\cite{PhysRevD.71.123001};
   \cite{Daniel.Stern_2010};\cite{M.Moresco_2012};\cite{Moresco_2016};
   \cite{10.1093/mnras/stx301};\cite{10.1093/mnrasl/slv037};\cite{Borghi_2022};
   \cite{Jiao_2023}),the Pantheon+ set of 1701 SNe Ia (\cite{Scolnic_2022}),
   the BAO data from SDSS and DESI 2024.

@@

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_{\text{data}}$ and constrains $H_0$
   in advance. To eliminate the influence of $M_b$, we give the $\chi^2$
   as \begin{equation}
      \chi_{\text{SNe}}^2=A-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $A=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})^2/\sigma_i^2,
   B=\sum_{i=1}^{1701}(\mu_{\text{th}}-\mu_{\text{data}})/\sigma_i^2,C=\sum_i^{1701}1/\sigma_i^2$,
   the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.
   \end{equation}
   However, the Eq.(\ref{eq:9}) just need the $H_0$ to caculate the $M_x$
   and we would still use the simple likelihood function as
   \begin{equation}
      \tilde{\chi}_{\text{SNe}}^2=\Delta^{\text{T}}C^{-1}\Delta,
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data.
   $B=\Delta^{\text{T}}C^{-1}$ and $C=\text{sum}(C^{-1})$.

@@

   The $L_X-L_{UV}$ relation of quasar is usually written as
   \begin{equation}
      L_X=\beta+\gamma\log L_{UV},
   \end{equation}
   which gives that
   \begin{equation}
      \log_{10}(F_X)=\beta+(\gamma-1)\log_{10}(4\pi)+\gamma\log_{10}(F_{UV})
                    +2(\gamma-1)\log_{10}(d_L),
   \end{equation}
   where $d_L$ is the luminosity distance same as which in the SNe Ia,Eq.15
   so the $\chi^2$ function for the QSO data can be defined as
   \begin{equation}
      \chi_{\text{QSO}}^2=\sum_i^{2421}\frac{(y_{\text{th}}^2-y_{\text{data}}^2)}{s_i^2}
      -\ln(2\pi s_i^2),
   \end{equation}
   where $s_i^2=\mathrm{d}y_i^2+\gamma^2\mathrm{d}x_i^2+\delta^2$ refers to the 
   uncertainties on the $x_i\ (\log_{10}F_X)$ and $y_i\ (\log_{10}F_{UV})$ and
   $\delta_i$ reprensent the instrinsic dispersion.

@@

   \begin{table}[htbp]
      \caption{The BAO-only dataset from SDSS}
      \centering
      \begin{tabular}{lllll}
         \hline\hline
            $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
            \hline
            0.15 & & & 4.47$\pm$0.17 \\
            0.38 & 10.23$\pm$0.17 & 25$\pm$0.76 & \\
            0.51 & 13.36$\pm$0.21 & 22.33$\pm$0.58 & \\
            0.7 & 17.86$\pm$0.33 & 19.33$\pm$0.53 & \\
            0.85 & & & $18.33_{-0.62}^{+0.57}$ \\
            1.48 & 30.69$\pm$0.8 & 13.26$\pm$0.55 & \\
            2.33 & 37.6$\pm$1.9 & 8.93$\pm$0.28 & \\
            2.33 & 37.3$\pm$1.7 & 9.08$\pm$0.34 & \\
            \hline
      \end{tabular}
      \label{tab:2}
   \end{table}

   \begin{table}[htbp]
      \caption{The BAO dataset from DESI 2024}
      \centering
      \begin{tabular}{lllll}
         \hline\hline
            $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
            \hline
            0.295 & & & 7.93$\pm$0.15 \\
            0.51 & 13.62$\pm$0.25 & 20.98$\pm$0.61 & \\
            0.706 & 16.85$\pm$0.32 & 20.08$\pm$0.6 & \\
            0.93 & 21.71$\pm$0.28 & 17.88$\pm$0.35 & \\
            1.317 & 27.79$\pm$0.69 & 13.82$\pm$0.42 & \\
            1.491 & & & 26.07$\pm$0.67 \\
            2.33 & 39.71$\pm$0.94 & 8.52$\pm$0.17 & \\
            \hline
      \end{tabular}
      \label{tab:3}
   \end{table}

@@

>>>>

\subsection{Model 1: relation to the ΛCDM model}

@@

   To constrain the relevant IDM models (\cite{refId0}), we use the newly revised
   observational $H(z)$ data (OHD)(\cite{PhysRevD.71.123001};
   \cite{Daniel.Stern_2010};\cite{M.Moresco_2012};\cite{Zhang_2014};
   \cite{Moresco_2016};\cite{10.1093/mnras/stx301};\cite{10.1093/mnrasl/slv037};
   \cite{Borghi_2022};\cite{Jiao_2023}),the Pantheon+ set of 1701 SNe Ia 
   (\cite{Scolnic_2022}), the QSO data from \cite{Lusso_2020}, 
   the BAO data from SDSS and DESI 2024, and the reduced CMB parameters
   from Planck 2018.

@@

   We use the fiducial SN Ia magnitude ($M_b$) determined from SH0ES 2021 Cepheid 
   host distances(\cite{Riess_2022}), which gives the $\mu_{\text{data}}$ and we 
   give the $\chi^2$ as 
   \begin{equation}
      \chi_{\text{SNe}}^2=\Delta^{\text{T}}C^{-1}\Delta,
   \end{equation}
   where $\Delta=(\mu_{\text{th}}-\mu_{\text{data}})$ and $C^{-1}$ is the inverse of the
   covariance matrix of the SNe Ia data, the distance modulus is $\mu=5\log_{10}(d_L/\text{Mpc})+25$, and the
   luminosity distance $d_L$ can be given as a function of redshift $z$
   \begin{equation}
      d_L=(1+z)\int_0^z\frac{c\mathrm{d}z'}{H(z')}.\label{eq:15}
   \end{equation}
   To eliminate the advanced restriction to $H_0$ from $M_b$, we adopt
   the likelihood function as
   \begin{equation}
      \widetilde{\chi}_{\text{SNe}}^2=A-\frac{B^2}{C}+\ln\left(\frac{C}{2\pi}\right),
   \end{equation}
   where $A=\Delta^{\text{T}}C^{-1}\Delta,B=\Delta^{\text{T}}C^{-1}$ and $C$ is the sum of $C^{-1}$.
   Apparently these two functions give the same constraints in $\Omega_{2,0}$ and
   $\log_{10}(\kappa C_1)$

@@

   The $L_X-L_{UV}$ relation of quasar is usually written as
   \begin{equation}
      L_X=\beta+\gamma\log L_{UV},
   \end{equation}
   which gives that
   \begin{eqnarray}
      \nonumber\log_{10}(F_X)&=\beta+(\gamma-1)\log_{10}(4\pi)+\gamma\log_{10}(F_{UV})
                    \\ &+2(\gamma-1)\log_{10}(d_L),
   \end{eqnarray}
   where $d_L$ is the luminosity distance same as Eq.(\ref{eq:15}).
   So the $\chi^2$ function for the QSO data can be defined as
   \begin{equation}
      \chi_{\text{QSO}}^2=\sum_i^{2421}\frac{(y_{\text{th}}^2-y_{\text{data}}^2)}{s_i^2}
      -\ln(2\pi s_i^2),
   \end{equation}
   where $s_i^2=\mathrm{d}y_i^2+\gamma^2\mathrm{d}x_i^2+\delta^2$ refers to the 
   uncertainties on the $x_i\ (\log_{10}F_X)$ and $y_i\ (\log_{10}F_{UV})$ and
   $\delta_i$ reprensent the instrinsic dispersion.

@@

   \begin{table}[htbp]
      \caption{The BAO-only dataset from SDSS}
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
         \hline
         0.15 & & & 4.47$\pm$0.17 \\
         0.38 & 10.23$\pm$0.17 & 25$\pm$0.76 & \\
         0.51 & 13.36$\pm$0.21 & 22.33$\pm$0.58 & \\
         0.7 & 17.86$\pm$0.33 & 19.33$\pm$0.53 & \\
         0.85 & & & $18.33_{-0.62}^{+0.57}$ \\
         1.48 & 30.69$\pm$0.8 & 13.26$\pm$0.55 & \\
         2.33 & 37.6$\pm$1.9 & 8.93$\pm$0.28 & \\
         2.33 & 37.3$\pm$1.7 & 9.08$\pm$0.34 & \\
         \hline
      \end{tabular}
      \label{tab:2}
   \end{table}

   \begin{table}[htbp]
      \caption{The BAO dataset from DESI 2024}
      \centering
      \begin{tabular}{llll}
         \hline\hline
         $z_{\text{eff}}$ & $D_{\text{M}}/r_{\text{d}}$ & $D_{\text{H}}/r_{\text{d}}$ & $D_{\text{V}}/r_{\text{d}}$ \\
         \hline
         0.295 & & & 7.93$\pm$0.15 \\
         0.51 & 13.62$\pm$0.25 & 20.98$\pm$0.61 & \\
         0.706 & 16.85$\pm$0.32 & 20.08$\pm$0.6 & \\
         0.93 & 21.71$\pm$0.28 & 17.88$\pm$0.35 & \\
         1.317 & 27.79$\pm$0.69 & 13.82$\pm$0.42 & \\
         1.491 & & & 26.07$\pm$0.67 \\
         2.33 & 39.71$\pm$0.94 & 8.52$\pm$0.17 & \\
         \hline
      \end{tabular}
      \label{tab:3}
   \end{table}

@@

   We use the Markov chain Monte Carlo (MCMC) method based 
   on the opening package \texttt{emcee} to give a global constraints
   to the free parameters $\Omega_{2,0}$ and $\log_{10}(\kappa C_1)$ in 
   Model 1 and $n$ in Model 2.
   Besides, we also add the parameter $H_0$ or $h$ to give the constraints
   to the dark matter particle mass $M_x$.

\subsection{Model 1: mimicking the ΛCDM model}

   The Eq.(\ref{eq:8}) convergant to the flat ΛCDM model as
   $\log_{10}(\kappa C_1)\to-\infty$, therefore the constraints
   could only give the upper limit of what, and give the lower limit
   of $M_x$, respectively.

   \begin{figure}[htbp]
      \centering
   \end{figure}
